---
title: "SSMtoAtlantis"
output: html_document
Author: Alaia Morell
Date: 15/08/2025
---
# From SSM grid (FVCOM like) to regular grid (ROMS like)
## Load the environment 
```{r}
# devtools::install_github("alaiam/SSMtoAtlantis")
library(reticulate)
library(tidyverse)
library(tidync)
library(doParallel)
library(SSMtoAtlantis)
library(ncdf4)
# setup_python_env(env_name = "Salish_sea_env",
#    conda_path = "/home/atlantis/anaconda3/bin")
use_condaenv("Salish_sea_env")
year = 2011
scenario = "cc_nutrient_paper"
filename_hyd = "/nfsdata/NandCCpaper_Dec2025/N_T_control/yr2011_hyd_time_avg_crop.nc"
filename_wqm = "/nfsdata/NandCCpaper_Dec2025/N_T_control/yr2011 _wqm_time_avg_crop.nc"
select <- dplyr::select #TODO: fix step B for current --> the function does not run without this row


```



```{r}
SSMtoAtlantis::preprocess()
?preprocess
?face_horizontal_equivalence
?setup_python_env
?StepA
?StepB
?join_dailyfiles


```


## Physical variables
```{r}

filename = filename_hyd # mandatory step for now
SSMtoAtlantis::StepA(year, "temperatureee", scenario, filename = filename)
SSMtoAtlantis::StepA(year, "U" ,          scenario, filename = filename)

```
## Water quality variables

```{r}
filename = filename_wqm # mandatory step for now
SSMtoAtlantis::StepA(year, "SP",     scenario, filename = filename)
SSMtoAtlantis::StepA(year, "SZ",     scenario, filename = filename)
SSMtoAtlantis::StepA(year, "LPON",   scenario, filename = filename)
SSMtoAtlantis::StepA(year, "POM_sed",   scenario, filename = filename)
SSMtoAtlantis::StepA(year, "RDON",   scenario, filename = filename)
SSMtoAtlantis::StepA(year, "NO3",    scenario, filename = filename)
SSMtoAtlantis::StepA(year, "O2", scenario, filename = filename)
```

# From regular grid to Atlantis grid (irregular polygons) daily files

## Physical variables
# Step B crashes without 32 Go for the RAM

```{r}
SSMtoAtlantis::StepB(year, "U", scenario) # Very heavy for your computer: be careful before to run
SSMtoAtlantis::StepB(year, "temperatureee", scenario)
```

## Water quality variables
# Step B crashes without 32 Go for the RAM
```{r}
SSMtoAtlantis::StepB(year, "SP", scenario)
SSMtoAtlantis::StepB(year, "SZ", scenario)
SSMtoAtlantis::StepB(year, "LPON", scenario)
SSMtoAtlantis::StepB(year, "RDON", scenario)
SSMtoAtlantis::StepB(year, "NO3", scenario)
SSMtoAtlantis::StepB(year, "O2", scenario)
```

# From Atlantis grid (irregular polygons) daily files to yearly files

```{r}
SSMtoAtlantis::join_dailyfiles(year, "temperature", scenario)
SSMtoAtlantis::join_dailyfiles(year, "U", scenario)
SSMtoAtlantis::join_dailyfiles(year, "SP", scenario)
SSMtoAtlantis::join_dailyfiles(year, "SZ", scenario)
SSMtoAtlantis::join_dailyfiles(year, "LPON", scenario)
SSMtoAtlantis::join_dailyfiles(year, "RDON", scenario)
SSMtoAtlantis::join_dailyfiles(year, "NO3", scenario)
SSMtoAtlantis::join_dailyfiles(year, "O2", scenario)
```


# Example of how to create an Atlantis input for a time serie
```{r}

# Officially from 1981 to 2011, using looped 2011 for the entire period (burn-in period)
initialization.period = SSMtoAtlantis::linear_interpolation(year_start = 1981, year_end = 2011, 
                                              year_name_decade_start = 2011, 
                                              year_name_decade_end =2011, 
                                              variable = "temperature", scenario = "status_quo")

# From 2011 to 2025, using year 2011 and decade 2020s
Y2010s <- SSMtoAtlantis::linear_interpolation(year_start = 2011, year_end = 2025, 
                                              year_name_decade_start = 2011, 
                                              year_name_decade_end =2020, 
                                              variable = "temperature", scenario = "status_quo")

# From 2011 to 2025, using year 2011 and decade 2020s
Y2020s <- SSMtoAtlantis::linear_interpolation(year_start = 2025, year_end = 2035, 
                                              year_name_decade_start = 2020, 
                                              year_name_decade_end =2030, 
                                              variable = "temperature", scenario = "status_quo")

# From 2011 to 2025, using year 2011 and decade 2020s
Y2030s <- SSMtoAtlantis::linear_interpolation(year_start = 2035, year_end = 2045,
                                              year_name_decade_start = 2030, 
                                              year_name_decade_end =2040,  
                                              variable = "temperature", scenario = "status_quo")

# Combining the full time serie
full.time.serie <- abind::abind(initialization.period, Y2010s, Y2020s, Y2030s, along = 3)

# Plot
plot(x = (1:46720)/730+1981, y = apply(X = full.time.serie[1,2:10,], FUN = mean, MARGIN = 2), 
     xlab = "Year", ylab = "Temperature")

# Write the nc.files
write_interpoled_array(full.time.serie, start_year = 1981, end_year = 2045, scenario = "status_quo", 
                                      variable = "temperature")

```



